\include{preamble}

\title{Infinite Alphabet Parsing \\ and \\ Compact LR Parser}
\date{July 30, 2017}
\author{Alexei A. Smekalkine <ikle@ikle.ru>}

\begin{document}

\maketitle

\begin{abstract}
This article describes generalization of well-known shift-reduce
parsing techniques to infinite input alphabet and compact LR parser
generator based on this generalization.
\end{abstract}

\section{Base Definitions}
\def\A{\mathcal{A}}
\def\G{\mathcal{G}}
\def\P{\mathcal{P}}

\begin{enumerate}
\item	Let the alphabet $\A$ be a possibly infinite set of terminal
	symbols. We will use lower case Latin letters represented by a
	normal font to denote terminals from other symbols.

\item	Let the generator grammar $\G$ be a non-empty set of generator
	rules of the form $A \rightarrow \P$ where:
	\begin{itemize}
	\item	$A$ is non-terminal symbol. We will use upper case
		Latin letters represented by a normal font to denote
		non-terminals.

	\item	$\P$ is the production of rule defined below.
	\end{itemize}

\item	Let $\alpha$ be a possibly empty sequence of symbols, terminals
	and non-terminals, and $|\alpha|$ be the lenght of this sequence.
	We will use lower case Greek letters represented by a normal
	font to denote such sequences.

\item	Let $\P = \alpha$ be a production of a rule.
\end{enumerate}

\section{Look-Ahead}
\def\L{\mathcal{L}}
\def\F{\mathcal{F}}
\def\e{\varepsilon}
\def\J{\mathcal{J}}

The look ahead is a possibly empty sequence of input symbols (terminals)
that used together with DFA state to resolve conflicts with next state
decision while parsing.

\begin{enumerate}

\item	Let $\L_n$ be look-ahead set of sequences of length upto $n$.

\item	Let $\F_n(\alpha)$ be a set of a prefixes of length upto $n$ of
	sequences of terminals that can be derived from production
	$\alpha$. The special marker $\e$ denotes empty prefix, $|\e| = 0$.

	$$
	\F_0(\alpha) = \emptyset
	$$

	$$
	\F_1(\alpha) =
	\begin{cases}
		\{ \e \},	& \text{if } |\alpha| = 0		\\
		\{ a  \},	& \text{if } \alpha = a \beta		\\
		\F_1(B),	& \text{if }
			\alpha = B \beta \wedge \e \not \in \F_1(B)	\\
		\F_1(B) \cup \F_n(\beta),
				& \text{otherwise}
	\end{cases}
	$$

	where

	$$
	\F_n(B) = \bigcup_{\forall i \wedge A_i = B} \F_n(P_i)
	$$

	Note, we omit now general case for $n > 1$ for simplicity.
%	this will be defined later.
\end{enumerate}

\section{Item Sets}
\def\I{\mathcal{I}}
\def\S{\mathcal{S}}
\def\G{\mathcal{G}}

\begin{enumerate}
\item   Let $\I_{n,i,j,k}$ be the item number $j$ of the rule
	$A_i \rightarrow \P_i$ augmented with a look ahead set
	$\L_n(k,A_i)$ for the state $k$:

	$$
	\I_{n,i,j,k} = [ A_i \rightarrow \alpha \bullet \beta, \L_n(k,A_i) ]
	$$

	where $\alpha \beta = \P_i$, $|\alpha| = j$. We can read $j$ as
	the position of a marker $\bullet$ in the production $\P_i$.

\item	Let $\S_k(B,\G)$ be the start set of items for not-terminal $B$
	on the generator look ahead set $\G$ for state number $k$:

	$$
	\S_k(B,\G) = \{ \I_{i,0,k} | \forall i \wedge A_i = B \}
	$$

	where

	$$
	\L_0(k,B) = \emptyset
	$$

	$$
	\L_1(k,B) =
	\begin{cases}
		\F_1(B), & \text{if } \e \not \in \F_1(B) \\
		\F_1(B) \setminus \{ \e \} \cup \G, & \text{otherwise}
	\end{cases}
	$$

\item	Let $\I_{n,i,j,k}^*(\G)$ be the closure of the item $\I_{n,i,j,k}$
	on the generator look ahead set $\G$:

	$$
	\I_{n,i,j,k}^*(\G) =
	\begin{cases}
	\{ \I_{n,i,j,k} \} \cup \S_k(B,\G)^* \text, &
		\text{if } P_i = \alpha \bullet B \beta \\
	\{ \I_{n,i,j,k} \} \text, & \text{othewise}
	\end{cases}
	$$

	where

	$$
	\S_k(B,\G)^* = \bigcup_{\I \in \S_k(B,\G)} \I^*(\G)
	$$
\end{enumerate}

\section{State}
\def\C{\mathcal{C}}
\def\eoi{\text{\$}}

\begin{enumerate}
\item	Let the core $\C_k$ for the state number $k$ be a set of items.

\item	Let $\C_k^*$ be the closure of the core $\C_k$

	$$
	\C_k^* = \bigcup_{\I \in C_k} \I^*
	$$

\item	Let start state number $0$ be identified by $\C_0^*$ where

	$$
	\C_0 = \I_{n,0,0,0} \text{ and }
	\L_n(0,A_0) = \{ \eoi \} \text{ for } n > 0
	$$

	where $\eoi$ is a special end-of-input terminal.
\end{enumerate}

\iffalse
\section{draft}
	blalba: and for $k > 1$

	$$
	\F_k(\alpha) =
	\begin{cases}
		\J_k(B, \beta),	& \text{if }
			\alpha = B \beta \wedge \e \not \in \F_1(B)	\\
		\J_k(B, \beta) \cup \F_k(\beta),
				& \text{if }
			\alpha = B \beta \wedge \e      \in \F_1(B)
	\end{cases}
	$$

	where

	$$
	\J_k(B, \beta) = \{ a \alpha | a \in \F_1(B) \wedge \alpha \in \F_{k-1}(\beta) \}
	$$

	blabla:
	$$
	\F_k(B) = \bigcup_{\forall A_i = B \wedge \{ l | l \in \L(state) \}}
		\F_k(\P_i || l)
	$$
\fi

\end{document}
